---
title: "Soil invertebrates"
description: |
  Land use effects on soil meso- and macro-fauna communities
output:
    distill::distill_article:
      toc: true
      toc_depth: 3
---

I tested soil We have five different land use types (pasture, silvopasture, secondary forest, "old growth", and teak plantation) and two sampling locations (Eco Venao and Achotines/Vernes Land). If we were to replicate this study we'd want more sampling locations and more replicates per location. As it is each sample from each land use type at each location is a composite of three sub-samples collected 5m apart and homogenized in the field, in an attempt to avoid micro-climate variability. 

Here is a .csv file with raw data for the invert pilot study:

```{r echo=FALSE}
xfun::embed_file('nov22_soil.csv')
```  

<p>&nbsp;</p>

First I cleaned up the data and prepared it for analysis, then calculated diversity indexes: 

<details markdown="1">
<summary><strong>Click here</strong> for setup information.</summary>

```{r setup, message=FALSE, results = 'hide'}
setwd("C:/Users/Hubert/Desktop/Earthshot/Nov 2022 markdown/Nov22_pilot")

data <- read.csv("nov22_soil.csv")

#load packages
library(vegan)
library(ggplot2)
#library("dplyr")
library(tidyr)
library(viridis)
library(gridExtra)
library('funrar')

## Initial cleanup, organization

#make dates recognizable by R
data$date<-as.Date(data$date, tryFormats = c("%Y-%m-%d", "%Y/%m/%d"))
                   

#remove missing count values
data<-data[which(complete.cases(data$count)),]

#create dataset with unknowns (for total abundance)
data.unknown<-data

#create dataset without unknowns (for diversity and community composition)
data <- subset(data, !taxon == "Unknown",)

#make my data wide for vegan package
wide.data <- data %>%
  spread(key=taxon,
         value=count)

#make same wide dataframes for data with unknowns
wide.data.unknown <- data.unknown %>%
  spread(key=taxon,
         value=count)

#Splitting data frame into count and environmental data
wide.data2<-wide.data[,-c(1:12)]     
wide.data3<-wide.data[,-c(13:65)]
wide.data.unknown2<-wide.data.unknown[,-c(1:12)]
wide.data.unknown3<-wide.data.unknown[,-c(13:66)]

### during the real sampling we'll collect soil moisture values so that we can mass correct count data to dry mass. For now I'm pretending it's all 40% 

#converting to count/1000g
wide.data2 <- (1000/wide.data3$calc.dry.mass)*wide.data2
wide.data.unknown2 <- (1000/wide.data.unknown3$calc.dry.mass)*wide.data.unknown2

#Next I convert mass-corrected absolute abundance to relative abundance, which is used later for total abundance, individual taxon abundance, and community structure modelling. 
#Before the conversion I create data frames backing up the mass-corrected absolute count data for diversity index calculations. 

#Creating versions of the count data that are corrected for mass, but not made relative (for diversity indexes)
wide.data.absolute <- wide.data2
wide.data.unknown.absolute <- wide.data.unknown2

#reincorporating, and duplicating mass-corrected absolute count data (only doing it for unknowns because this data frame will be used for abundance)
wide.data.unknown[,c(13:66)] <- wide.data.unknown.absolute
wide.data.unknown.absolute2 <- wide.data.unknown

#converting absolute to relative abundance
wide.data2 <- as.matrix(wide.data2)
wide.data.unknown2 <- as.matrix(wide.data.unknown2)
wide.data2 <- make_relative(wide.data2)
wide.data.unknown2 <- make_relative(wide.data.unknown2)
wide.data2 <- as.data.frame(wide.data2)
wide.data.unknown2 <- as.data.frame(wide.data.unknown2)

#reincorporating mass-corrected relative count data into original data frame
wide.data[,c(13:65)] <- wide.data2
wide.data.unknown[,c(13:66)] <- wide.data.unknown2


#Finally I use the ['vegan'](https://cran.r-project.org/web/packages/vegan/index.html) package calculate richness, diversity, evenness, and total abundance, and make a summary table of these values.

taxa <- specnumber(wide.data.absolute)                  #Tallying number of taxa
shannon<-diversity(wide.data.absolute)                  #Shannon diversity index
simpson<- diversity(wide.data.absolute, "simpson")      #Simpson diversity index
J<-shannon/log(taxa)                                    #Pielou's evenness
abundance<-rowSums(wide.data.unknown.absolute)          #total absolute abundance per 1000g (including unknowns)

data.summary<-data.frame(wide.data3,taxa, abundance, J, shannon, simpson)
```

</details>

<p>&nbsp;</p>

### Total abundance 

```{r, echo=FALSE, message = FALSE, warning = FALSE}
land.use.order <- c('"old-growth"', 'secondary', 'silvopasture', 'pasture','teak')

##### Abundance
ggplot(data=data.summary, aes(x=land.use, y=abundance, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), aes(x = factor(land.use, level = land.use.order)), width = .5)+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Total abundance of soil fauna", 
       x="Land Use Type", y = "Abundance (count/1kg soil")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5))
```


### Taxonomic Richness

```{r, echo=FALSE, message = FALSE, warning = FALSE}
ggplot(data=data.summary, aes(x=land.use, y=taxa, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Taxonomic richness of soil fauna", 
       x="Land Use Type", y = "Ordinal taxonomic richness")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5))
```

### Shannon Diversity

```{r, echo=FALSE, message = FALSE, warning = FALSE}
ggplot(data=data.summary, aes(x=land.use, y=shannon, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Shannon diversity of soil fauna", 
       x="Land Use Type", y = "Shannon index")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5))
```

### Simpson Diversity

```{r, echo=FALSE, message = FALSE, warning = FALSE}
ggplot(data=data.summary, aes(x=land.use, y=simpson, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="simpson diversity of soil fauna", 
       x="Land Use Type", y = "Simpson index")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5))
```

## Responses of abundant taxa

#### Hymenoptera
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Hymenoptera absolute
hym.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Hymenoptera, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Hymenoptera absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))


##### Hymenoptera relative
hym.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Hymenoptera, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Hymenoptera relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(hym.ab, hym.rel, ncol=2, widths = c(1.65,2))
```

#### Coleoptera
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Coleoptera absolute
col.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Coleoptera, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Coleoptera absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))

##### Coleoptera relative
col.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Coleoptera, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Coleoptera relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(col.ab, col.rel, ncol=2, widths = c(1.65,2))
```

#### Hemiptera
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Hemiptera absolute
hem.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Hemiptera, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Hemiptera absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))


##### Hemiptera relative
hem.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Hemiptera, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Hemiptera relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(hem.ab, hem.rel, ncol=2, widths = c(1.65,2))
```

#### Protura
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Protura absolute
pro.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Protura, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Protura absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))

##### Protura relative
pro.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Protura, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Protura relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(pro.ab, pro.rel, ncol=2, widths = c(1.65,2))
```

#### Collembola
```{r, echo=FALSE, message = FALSE, warning = FALSE}
wide.data.unknown$Collembola <- wide.data.unknown$Entomobryomorpha + wide.data.unknown$Poduromorpha + wide.data.unknown$Symphypleona
wide.data.unknown.absolute2$Collembola <- wide.data.unknown.absolute2$Entomobryomorpha + wide.data.unknown.absolute2$Poduromorpha + wide.data.unknown.absolute2$Symphypleona

##### Collembola absolute
coll.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Collembola, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Collembola absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))

##### Collembola relative
coll.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Collembola, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Collembola relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5))

grid.arrange(coll.ab, coll.rel, ncol=2, widths = c(1.65,2)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

```

#### Oribatida
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Oribatida absolute
ori.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Oribatida, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Oribatida absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))

##### Oribatida relative
ori.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Oribatida, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Oribatida relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(ori.ab, ori.rel, ncol=2, widths = c(1.65,2))
```

#### Other Acari
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### other.acari absolute
aca.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=other.acari, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Other Acari absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))


##### other.acari relative
aca.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=other.acari, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="other.acari relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(aca.ab, aca.rel, ncol=2, widths = c(1.65,2))
```

#### Pseudoscorpiones
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Pseudoscorpiones absolute
pse.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Pseudoscorpiones, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Pseudoscorpiones absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))


##### Pseudoscorpiones relative
pse.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Pseudoscorpiones, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Pseudoscorpiones relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(pse.ab, pse.rel, ncol=2, widths = c(1.65,2))
```

#### Diplopoda
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Diplopoda absolute
dip.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Diplopoda, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Diplopoda absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))

##### Diplopoda relative
dip.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Diplopoda, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Diplopoda relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(dip.ab, dip.rel, ncol=2, widths = c(1.65,2))
```

#### Symphyla
```{r, echo=FALSE, message = FALSE, warning = FALSE}
##### Symphyla absolute
sym.ab <- ggplot(data=wide.data.unknown.absolute2, aes(x=land.use, y=Symphyla, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Symphyla absolute", 
       x="Land Use Type", y = "Abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position="none") + scale_x_discrete(guide = guide_axis(n.dodge = 2))


##### Symphyla relative
sym.rel <- ggplot(data=wide.data.unknown, aes(x=land.use, y=Symphyla, fill=site)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.50, aes(x = factor(land.use, level = land.use.order)))+
  #scale_fill_brewer(palette="Paired")+ 
  labs(title="Symphyla relative", 
       x="Land Use Type", y = "Relative abundance (count/1kg soil)")+
  theme(text=element_text(family="Arial")) + 
  scale_fill_viridis(discrete=TRUE, option="viridis", begin = .3,end = .75,)+
  theme_minimal()+ theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(guide = guide_axis(n.dodge = 2))

grid.arrange(sym.ab, sym.rel, ncol=2, widths = c(1.65,2))

```